---
title: "Statistical Inference Project"
author: "S. Smolenski"
---

# Overview:
In this project, I investigate the exponential distribution in R and compare it to the Central Limit Theorem. I consider the distribution of averages of 40 exponentials, performing 1000 simulations to create the data used for this project.

# Simulations:
I set the parameters for the exponential function and perform the simulation.

```{r}
n=40
numsim=1000
lambda<-.2
Sim<-matrix(rexp(numsim*n,lambda), numsim)
Means<-apply(Sim, 1, mean)

```

# Sample Mean versus Theoretical Mean:
The means are approximately the same:

## Sample mean:
```{r}
SimMn<-mean(Means)
SimMn
```

## Theoretical Mean:
```{r}
TheoMn<-1/lambda
TheoMn
```

## Visual Comparison of Means:
```{r}
hist(Means, col="thistle", main="Simulated Means")
abline(v=SimMn, col="dodgerblue4", lwd=2)
abline(v=TheoMn, col="darkmagenta", lty=3)
legend("topright", lwd=c(3,1), lty=c(1,2), col=c("dodgerblue4","darkmagenta"), legend=c("Simulated Mean", "Theoretical Mean"))
```

# Sample Variance versus Theoretical Variance:

```{r}
SimVar<-var(Means)
SimVar
```

```{r}
TheoVar<-((-1/lambda)^2)/n
TheoVar
```

The theoretical variance is slightly lower than the simulated variance, implying that the simulated distribution will have a slightly larger spread. The larger numsim becomes, the closer to the theoretical variance this value should be.

# Distribution

If we overlay a normal density function on top of our simulated data, we find that the distribution is roughly normal:

```{r}
hist(Means, col="cadetblue", main="Simulated Means")
curve( dnorm(x, mean=SimMn,sd=sqrt(SimVar))*500, 2,8, add=TRUE, col="gray42")
```
