---
title: "Statistical Inference Project, Part 2"
author: "S. Smolenski"
---

# Introduction
This project explores the ToothGrowth R dataset, which describes: "The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC)."

# Exploratory Analysis

Loading the dataset:

```{r}
data(ToothGrowth)
str(ToothGrowth)
```

Plotting Vitamin C dose versus growth length. Colors are used to separate delivery methods.

```{r}
with(ToothGrowth, plot(dose,len,col=supp, pch=20, main="Vitamin C Dose vs \n Tooth Growth \n in Guinea Pigs", xlab="Vitamin C Dose (mg/day)", ylab="Length of Odontoblasts"))
```

```{r}
with(ToothGrowth, boxplot(len~supp*dose,col=c("thistle","cadetblue"), main="Vitamin C Dose vs \n Tooth Growth \n in Guinea Pigs", xlab="Vitamin C Dose (mg/day)", ylab="Length of Odontoblasts"))
```

# Summary
The data seem to imply that tooth growth is correlated with Vitamin C dose, though the relationship between delivery method is less clear. For .5 and 1.0 mg/day, the OJ supplement wins out over the VC supplement, but at 2.0 mg/day, they seem roughly equivalent.

# Confidence Interval Testing
Let us more closely examine the relationship between supplement type and tooth growth, since this is the more ambiguous relationship.

We perform a two-sided t test comparing the growth produced by OJ and VC, under the null hypothesis that there is no distinction between the two supplement types.

```{r}
OJ=ToothGrowth$len[ToothGrowth$supp=="OJ"]
VC=ToothGrowth$len[ToothGrowth$supp=="VC"]
t.test(OJ, VC, conf.level=.975)
```

# Conclusions
Under a two-sided, 95% confidence level, we obtain a P-value of .06. In other words, we do not reject the null hypothesis. We must conclude from this test that there is no significant difference between the two supplement types.

